---
title: "R for Data Science 2e - Chapter 03 - Data Transformation"
author: "Farooq Mahmud"
format: 
  html:
    self-contained: true
    echo: true
    message: false
    warning: false
    error: true # IMPORTANT NOTE - this makes the file render even if there is an error. it is imperative that you check your .html document *before submission* to make sure that it has all results in it.
editor: source
---

**1. In a single pipeline for each condition, find all flights that meet the condition:**

- **Had an arrival delay of two or more hours**
```{r}
library(nycflights13)
library(tidyverse)
library(dplyr)

flights |>
  filter(arr_delay >= 120) |>
  relocate(arr_delay, .before = year) |>
  head(3)
```

- **Flew to Houston (IAH or HOU)**
```{r}
flights |>
  filter(dest %in% c("IAH", "HOU")) |>
  relocate(dest, .before = year) |>
  head(3)
```

- **Were operated by United, American, or Delta**
```{r}
flights |>
  filter(carrier %in% c("UA", "AA", "DL")) |>
  relocate(carrier, .before = year) |>
  head(3)
```

- **Departed in summer (July, August, and September)**
```{r}
flights |>
  filter(month %in% c(7, 8, 9)) |>
  head(3)
```

- **Arrived more than two hours late, but didn’t leave late**
```{r}
flights |>
  filter(arr_delay >= 120) |>
  filter(dep_delay == 0) |>
  relocate(arr_delay, .before = year) |>
  relocate(dep_delay, .after = arr_delay) |>
  head(3)
```

- **Were delayed by at least an hour, but made up more than 30 minutes in flight**
```{r}
flights |>
  filter(dep_delay >= 60) |>
  filter(dep_delay - arr_delay > 30) |>
  relocate(dep_delay, .before = year) |>
  relocate(arr_delay, .after = dep_delay) |>
  head(3)
```

**2. Sort flights to find the flights with the longest departure delays. Find the flights that left earliest in the morning.**
```{r}
flights |>
  arrange(desc(dep_delay)) |>
  relocate(dep_delay, .before = year) |>
  head(3)
flights |>
  arrange(dep_time) |>
  relocate(dep_time, .before = year) |>
  head(3)
```

**3. Sort flights to find the fastest flights. (Hint: Try including a math calculation inside of your function.)**
```{r}
flights |>
  mutate(mph = (distance / air_time) * 60) |>
  arrange(desc(mph)) |>
  relocate(mph, .before = year) |>
  head(3)
```

**4. Was there a flight on every day of 2013?**
```{r}
daily_flights <- flights |>
  group_by(year, month, day) |>
  summarize(flights_per_day = n())
nrow(daily_flights) == 365
```

**5. Which flights traveled the farthest distance? Which traveled the least distance?**
```{r}
most_distance <- flights |>
  arrange(desc(distance)) |>
  relocate(distance, .before = year)
head(most_distance, 3)

least_distance <- flights |>
  arrange(distance) |>
  relocate(distance, .before = year) |>
  na.omit()
head(least_distance, 3)
```
